\emph{Gaia} DR2 adds clues for dynamical ejection of a $\sim$$150\,M_\odot$ star

### 2018-05-22 VFTS682.tex
tuned down abstract

\title{\emph{Gaia} DR2 identifies VFTS 682 as the most massive % , dynamically
                                % ejected %% to fit in one line
  runaway star known to date}

\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
  R.~P.~van~der~Marel$^{3,4}$,   I.~Platais$^{4}$, J. Bestenlehner$^{5}$,
   V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{7,8}$,  A.~de~Koter$^{1}$,
 , H.~Sana$^{9}$, F.~R.~Schneider$^{10}$, J.~S.~Vink$^{11}$}
\affil{$^{1}${\todo{check affil. num}Astronomical Institute Anton Pannekoek, University of
    Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
    $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
    E-28691 Villanueva de la Ca\~nada, Madrid, Spain} \\
       $^{3}$ {Space Telescope Science Institute, 3700 San Martin Drive,
       Baltimore, MD 21218, USA}\\
        $^{4}$ {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
      $^{}${Department of Physics and Astronomy, Hicks Building,
      Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
         $^{}$ {National Research Council, Herzberg Astronomy \& Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7, Canada} 
     $^{}$ {School of Astronomy \& Space Science, University of the Chinese
   Academy of Sciences, Beijing 100012, China}\\
    $^{}$ {National Astronomical Observatories, Chinese Academy of
      Sciences, Beijing 100012, China}\\
     $^{}$ {Department of Physics, University of Oxford, Keble Road,
       Oxford OX1 3RH, UK} \\
       $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}   
    $^{}$ {Space Telescope Science Institute, 3700 San Martin Drive,
       Baltimore, MD 21218, USA}\\
      {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
     $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
     $^{}$ {National Research Council, Herzberg Astronomy \& Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7, Canada}     
}

%\thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
\date{}
\begin{abstract}
The formation and fates of the most massive stars remains among the most elusive questions in stellar astrophysics. 
%
Previous spectroscopic studies identified VFTS 682 as one of the most
massive stars known. It resides in the field of the 30 Dor region in
the Large Magellanic Cloud at a projected distance of 29\,pc of the
star cluster Radcliffe 136 (R136). The lack of a detected offset in radial velocity\todo{can we really say that?RV
  is about 30kms...}
derived from the spectra, led to speculations whether such extreme
stars can form in relative isolation . 
%
\todo{add mention of HST}We use the unprecedented proper motions from the second \emph{Gaia} data
release (DR2) to investigate the kinematics of this object. We show
that VFTS 682 is a bonafide runaway star moving away from the star
cluster R136 with a relative
projected two-dimensional speed of about $\sim$$34\pm17\kms$, implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when
accounting for the velocity component along the line of sight. The
inferred age, distance, and the magnitude and direction of the
velocity are consistent with a dynamical ejection scenario from the cluster. 
%
This makes VFTS 682 the most massive runaway star known to date. It proves that star clusters are capable of ejecting their most massive members. We discuss the implications for the existence of stars with masses well above $100\Msun$ inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.

>>>>

% \title{\emph{Gaia} DR2 identifies VFTS 682 as the most massive % , dynamically
%                                 % ejected %% to fit in one line
%   runaway star known to date}
\title{\emph{Gaia} DR2 adds clues for dynamical ejection of
  a $\sim$$150\,M_\odot$ star}
\author{M.~Renzo$^{1}$, S.~E.~de~Mink$^{1}$, D.~J.~Lennon$^{2}$,
  R.~P.~van~der~Marel$^{3,4}$,   I.~Platais$^{4}$, J. Bestenlehner$^{5}$,
   V.~H\'enault-Brunet$^{6}$,  S.~Justham$^{7,8}$,  A.~de~Koter$^{1}$,
 , H.~Sana$^{9}$, F.~R.~Schneider$^{10}$, J.~S.~Vink$^{11}$}
\affil{$^{1}${\todo{check affil. num}Astronomical Institute Anton Pannekoek, University of
    Amsterdam, 1098 XH Amsterdam, The Netherlands} \\
    $^{2}$ {ESA, European Space Astronomy Centre, Apdo. de Correos 78,
    E-28691 Villanueva de la Ca\~nada, Madrid, Spain} \\
       $^{3}$ {Space Telescope Science Institute, 3700 San Martin Drive,
       Baltimore, MD 21218, USA}\\
        $^{4}$ {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
      $^{}${Department of Physics and Astronomy, Hicks Building,
      Hounsfield Road, University of Sheffield, Sheffield S3 7RH, UK}\\
         $^{}$ {National Research Council, Herzberg Astronomy \& Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7, Canada} 
     $^{}$ {School of Astronomy \& Space Science, University of the Chinese
   Academy of Sciences, Beijing 100012, China}\\
    $^{}$ {National Astronomical Observatories, Chinese Academy of
      Sciences, Beijing 100012, China}\\
     $^{}$ {Department of Physics, University of Oxford, Keble Road,
       Oxford OX1 3RH, UK} \\
       $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}   
    $^{}$ {Space Telescope Science Institute, 3700 San Martin Drive,
       Baltimore, MD 21218, USA}\\
      {Center for Astrophysical Sciences, Department of Physics \& Astronomy, Johns Hopkins University, Baltimore, MD 21218, USA}\\
     $^{}$ {Armagh Observatory, College Hill, Armagh BT61 9DG, UK}\\
     $^{}$ {National Research Council, Herzberg Astronomy \& Astrophysics, 5071 West Saanich Road, Victoria, BC, V9E 2E7, Canada}     
}

%\thanks{Corresponding author:  M.~Renzo, \href{mailto:m.renzo@uva.nl}{m.renzo@uva.nl}}
\date{}
\begin{abstract}
The formation and fates of the most massive stars remain among the most elusive questions in stellar astrophysics. 
%
Previous spectroscopic studies identified VFTS 682 as one of the most
massive stars known. It resides in the field of the 30 Dor region in
the Large Magellanic Cloud at a projected distance of 29\,pc of the
star cluster Radcliffe 136 (R136). The inconclusiveness of the radial
velocity analysis for this emission line star led to speculations whether such extreme
stars can form in relative isolation. 
%
We use the unprecedented proper motions from the second \emph{Gaia} data
release (DR2) together with multiepoch \emph{Hubble Space Telescope} proper motions to investigate the kinematics of this object. We show
that the projected velocity of VFTS 682 is consistent with it being a
runaway star dynamically ejected from R136. Its
projected two-dimensional speed relative to the cluster is $\sim$$34\pm17\kms$, implying a three-dimensional speed of about $\sim$$40\pm20\kms$ when
accounting for the (uncertain) velocity component along the line of sight. The
inferred kinematic age is also  consistent with the dynamical ejection scenario from the cluster. 
%
If confirmed by future astrometric data releases, this would make VFTS 682 the most massive runaway star known to date. It proves that star clusters are capable of ejecting their most massive members. We discuss the implications for the existence of stars with masses well above $100\Msun$ inside the cluster, the internal dynamics in dense star clusters and the final fate of this object.
